Debugging by \textit{lastChange}

### 2010-10-27 papers/pldi2011/sigplanconf-template.tex
reword abstract

\begin{abstract}
Backward search from bug symptoms to defects-which cause the bug-is a
natural way to debugging. However, due to the lack of support for
easily backward movement by traditional debugging approaches (e.g.,
breakpoint-based or log-based debugging), applying this strategy-if it
is not practically impossible-demands a lot of effort from
developers. Tracking origins to wrong values is a fundamental step in
this process and any aid in this regard can greatly reduce the
complexity of debugging for developers.

To address this problem, we propose a new functionality in debuggers,
\textit{lastChange}, which locates the last place a value is changed
and lets developer to collect data and ask for more
\textit{lastChange} queries at the located point. Although
\textit{lastChange} algorithm is based on program re-execution,
contrary to other replay-based algorithms which require exactly the
same re-executions, it only requires bug reproduction by
re-executions. In other words \textit{bug reproducibility} (i.e., a
test case is available which reproduces the bug) is the only
prerequisite for \textit{lastChange} functionality.

@@

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented it in an extension to the Firebug
JavaScript debugger[].%\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser[]. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface and
\textit{Querypointer} is developed over this interface.

@@

The Firefox JavaScript engine does not current support
\texttt{setPropertyChangeHook()} (see however \cite{bugzilla}).  For
the our \textit{lastChange} prototype we created a version of
\texttt{setPropertyChangeHook()} using only features of the current
engine.  The basic strategy is to get a reference to the object just
after its creation, then use \texttt{watch()} function to monitor
property changes in the object.  Setting a flag into the Firefox
JavaScript engine we can get file URL and line number for each object
creation (e.g., myFile.js, line 24). We set a breakpoint on this line
and analyze the source code to determine which object was created.

@@

An object might be created by one of these statements: An object
initializer(\texttt{\{...\}}) or \textit{new} operator (\texttt{new
  constructor()}) or function objects which can be created by
\texttt{function()} statement. By parsing the source code we can
easily recognize the statements create an object. The next step is
getting a reference to the new object.

The new object can be assigned to an object property or variable by
\texttt{=}. In these cases we keep the assignee statement at the left
side. The idea is that we create a list of assignee statements that
the new objects are assigned to them. We set a hook on the creation
line. Once the hook hits, we evaluate all statements. Then we do
stepping(step-over) and after each step we evaluate the
statements. Every statement which has a new value, we consider the new
value-if it is an object-as a new object. For example in Figure
~\ref{fig:objectCreation}(a), if the creation line number is 20, we
have only one statement which creates a new object and it is assigned
to \texttt{x.y}.

The new object can also be set as the property of a parent object by
\texttt{:}. This case is also treated similar to previous case. The
only difference is that the full path of property from the root parent
in the local scope should be considered as the assignee statement. For
example in Figure ~\ref{fig:objectCreation}(b), if the creation line
is 20, we keep \texttt{parentObject.newObject}. In this case stepping
should be continued until the end of line 22 for getting a reference
to the created new object.

There is another case for new objects and when they are passed as
arguments to other functions ~\ref{fig:objectCreation}(c). In these
cases instead of step-over we use step-in and we get the corresponding
argument as a new object.

%The algorithm is not complete and it is tried to support more and more cases

@@

@@

\subsection{\texttt{scopeId()} operation}
The scope id is kept as a variable with name \texttt{\_scopeId} in the
scope. We explained that we set a hook at the first instruction of the
function which defines the variable. Once this hook hits,
\texttt{\_scopeId} is set by calling JavaScript's dynamic compilation function \texttt{eval()}. For
example, executing \texttt{eval( var \_scopeId = 10)} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope.

\subsection{\texttt{setVariableChangeHook()} operation}
Only operations in the scope of these varibles can change their value; 
that includes the defining scope and scopes nested inside of them.
For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line variable is defined. In Figure
~\ref{fig:js-closure}, if the execution is paused at line 17 and the
last change of \texttt{var1} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different due
the fact that \texttt{var1} in \texttt{firstChild} is a local variable
but in \texttt{secondChild} is a closure variable.

@@

%---------------------------------------------------------------------------------------------------

\subsection{Execution reproduction}
Although execution reproduction is basicly should be provided by the
developer, we tried to devise some automatic way which reproduces the
execution. In \textit{Querypointer}, develooper can directly provide a
test case which reproduces the execution. It is also possible to use
two automatic reproduction option: record-replay or local-reproduction
.

Two elements should be carefully considered in execution
reproduction. First, the initial state should be the same as previous
execution. Second, the similar actions and events should be applied to
the program during the execution. In Record-replay approach, debugger
stores the initial page url and user actions and during replay phase,
opens the same url and simulates user actions. In the local
reproduction approach, debugger calls the funtion on the bottom of
stack with the same parameters. The idea behind local-reproduction is
that in many web pages it is enough to replay the event to reproduce
the bug. While record-replay approach gives better warranty about the
same initial states, the local-reproudction gives us shorter
cycles. It is also possible to uses thes approaches in
combination. For example, debugger locally reproduce the execution and
if it couldn't find the answer, reproduce the whole execution by
record-replay.
% should we talk about counting breakpoint hits for stopping at the right place?

\subsection{Data collection and presentation}
So far, we talked about the data collected at change events for
\textit{lastChange} algorithm. However, if developer needs to explore
the program state at a given point, more data is required. It is
possible to postpone data collection until user requests. It is also
possible to collect the list of available
variable. \textit{Querypointer} collects data in one level depth. More
data will be collected upon the developer's request by re-execution.

We needed to build an interface to different execution points which
are presented at the same time. The current \textit{Querypointer}
interface resembles the script panel interface while it keeps the
history of navigation.

@@

>>>>

\begin{abstract}

Developers often seek the origins of wrong values they see in their
debugger. Their search must be backwards in time: the code causing the
wrong value executed before the wrong value appeared. Searching with
breakpoint- or log- based debuggers demands persistance and
significant experience with the application being debugged. We propose
a new approach, an algorithm \textit{lastChange}, which automatically
locates the last point that a value was changed and related
information such as the call stack and program state. Starting from a
program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution. Thus
\textit{lastChange} combines the flexibility of breakpoint debugging
with the expressive power of log-based query debugging.  Contrary to
other replay-based algorithms which require exactly the same
re-executions, \textit{lastChange} only requires \textit{bug
  reproducibility}, a test case is available which reproduces the bug.

@@

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented it in an extension to the Firebug
JavaScript debugger\cite{Firebug}. %\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser\cite{Firefox}. %\cite{firefox-version-3.6}
The Firefox JavaScript engine provides a JavaScript debugging interface \cite{JSD} and
\textit{Querypointer} is developed over this interface.

@@

The Firefox JavaScript engine does not current support
\texttt{setPro-}
\texttt{pertyChangeHook()} (see however []).  For %\cite{bugzilla}
the our \textit{lastChange} prototype we created a version of
\texttt{setPropertyChangeHook()} using only features of the current
engine.  The basic strategy is to get a reference to the object just
after its creation, then use \texttt{watch()} function to monitor
property changes in the object.  Setting a flag into the Firefox
JavaScript engine we can get file URL and line number for each object
creation (e.g., myFile.js, line 24). We set a hook on this line
and analyze the source code to determine which object was created.

Here, we implicitly assumed that the object will be created at 
the same location in the next execution. So, what does happens if it
is not true? Well, once the execution reaches the reproduction point, it
reveals that the object has been created in different location. This time,
debugger re-executes the program considering both locations as possible 
object creation locations.

@@

An object might be created by one of these statements: object
initializer (\texttt{\{...\}}), \textit{new} operator (\texttt{new 
constructor()}) or function definition statment (\texttt{function()}). 
By parsing the source code we can easily recognize the statements 
create an object. The next step is getting a reference to the new object.

The new object can be assigned to an object property or variable by
an assignment (\texttt{=}). In these cases we keep the assignee statement 
at the left side. The idea is that we create a list of assignee statements 
that the new objects are assigned to them. We set a hook on the creation
line. Once the hook hits, we evaluate all assignee statements. Then 
we do stepping(step-over) and after each step we evaluate the
statements. Every statement which has a new value, we consider the new
value-if it is an object-as a new object. For example in Figure
~\ref{fig:objectCreation}(a), if the creation line number is 20, we
have only one statement which creates a new object and it is assigned
to \texttt{x.y}.

@@

(d) Case 4:
20 a[++i] = new myConstructor();

@@

The new object can also be set as the property of a parent object by a
colon (\texttt{:}). This case is also treated similar to previous case. The
only difference is that the full path of property from the root parent
in the local scope should be considered as the assignee statement. For
example in Figure~\ref{fig:objectCreation}(b), if the creation line
is 20, we keep \texttt{parentObject.newObject}. In this case stepping
should be continued until the end of line 22 for getting a reference
to the created new object.

There is another case for new objects and when they are passed as
arguments to other functions Figure~\ref{fig:objectCreation}(c). In these
cases instead of step-over we use step-in and we get the corresponding
argument as a new object.

Although this approach is successful in many ordinary cases, there are some 
cases that this approach doesn't work. For example in a case
where the new object is assigned to \texttt{a[++i]} where i is a variable
(Figure ~\ref{fig:objectCreation}(d)), \texttt{a[++i]} is kept in the list
of assignee statements. Obviously, evaluating this statement doesn't return
a reference to the new object.

\subsection{\texttt{scopeId()} operation}
Debugger sets a hook at the beginning of all code blocks need
a scope id whenever the block is loaded. The scope id is kept as a variable with name 
\texttt{\_scopeId} in the scope. Whenever the hook hits which means a new scope
is created, \texttt{\_scopeId} is set by calling JavaScript's dynamic compilation 
function \texttt{eval()}. For example, executing \texttt{eval( var \_scopeId = 10)} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope. \texttt{scopeId()}
operation returns the value of \texttt{\_scopeId} in the scope.

\subsection{\texttt{setVariableChangeHook()} operation}
Variables defined in local, closure and catch scopes are only change in
their scope; that includes the defining scope and scopes nested inside of them.
For example in Figure ~\ref{fig:js-closure}, variable \texttt{var1} in line 11 can only be
changed in lines 10 to 23. Therefore, after locating the function in
which the variable is defined, it is enough to parse the code inside
the function block and set a hook on all lines where the variable is
assigned a new value. 
We also set a hook at the first line of the function which is
corresponding to the line variable is defined. In Figure
~\ref{fig:js-closure}, if the execution is paused at line 17 and the
last change of \texttt{var1} is queried, two hooks on lines 16 and 17
will be set, but if the execution is paused at line 20 , four hooks on
lines 11, 12, 14, 20 will be set. These two cases are different due
the fact that \texttt{var1} in \texttt{firstChild} is a local variable
but in \texttt{secondChild} is a closure variable.

@@

\subsection{Re-Execution, reproduction point and data collection}
\textit{Querypointer} needs a test case to reproduce the
execution and conditions to correctly recognize the reproduction point. 
Although both elements can be directly provided by developer, \textit{Querypointer}
is also able to automatically create them from the first execution.

\textit{Querypointer} keeps track of breakpoint hits and user steppings. For example, 
if developer query \textit{lastChange} at the third hit of breakpoint \textit{b}, in
re-execution, third hit of breakpoint b recognized as the reproduction point.

To build a test case from one execution, two elements should be carefully 
considered. First, the initial state should be the same as the initial state
of execution. Second, similar actions and events should be applied to
the program during the execution. \textit{Querypointer} supports two mechanism
for automatic re-execution: local-reproduction, record-replay. The first one,
recalls the function at the bottom of call stack with the same parameters. The idea
behind this mechanism is that many bugs in web pages can be reproduced by firing
an event like clicking on a button. The second one, in the first execution, record phase, 
stores the initial page url and captures user actions. In the replay phase, it opens
the same url and applies user actions. The local reproduction mechanism provides
shorter re-execution cycles while the record-replay is more accurate about the
initial state.

In addition to the data collected at every change event for identifying the \textit{lastChange}
result, \textit{Querypointer} partially stores values in program state. If developer asks for 
some values which have not been stored, \textit{Querypointer} re-executes and collects the requested data.
There is a trade-off between the amount of data collected at every change event and the number of re-executions. 
By default, \textit{Querypointer} stores values up to one level depth in objects.

% Discuss what happens when assumptions changes, like object creation location.

@@

\bibitem[Firefox(2010)]{Firefox}
Firefox. \newblock http://www.mozilla.com.

\bibitem[Hofer(2006)]{Hofer}
C. Hofer, M. Denker, and C. St\'{e}phane. \newblock Implementing a backward-in-time debugger.
\newblock In Proceedings of\emph{NODe’06},
volume P-88, pages 17–32. Lecture Notes in Informatics, 2006.
